\documentclass[10pt, a4paper, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Implementing the \textit{Best move or first robot} change case}

\author{\IEEEauthorblockN{Matthew Eden}
\IEEEauthorblockA{\textit{Department of Electrical, Computer and Software
Engineering} \\
\textit{University of Auckland}\\
Auckland, New Zealand \\
mede607@aucklanduni.ac.nz}
}

\maketitle

\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
kalah, mancala, changeability, java
\end{IEEEkeywords}

\section{Planning the Change Case}
Before any code was actually written, a plan was devised describing the minimal
amount of changes that would be required to achieve the desired change case.
This plan is as follows.
\begin{enumerate}
  \item Find the \texttt{GameConfig} class.
  \item Change the value of \texttt{VERTICAL\_OUTPUT} from \texttt{true} to
    \texttt{false}.
  \item Find the \texttt{GameIO} class
  \item Create a new method: \texttt{public static int computeMove(Board board,
    int playerNum, IO io)}.
  \item Within this method, create a \texttt{List<Pit>} to contain the player's
    houses returned from
    \texttt{board.getPlayers().get(playerNum).getHouses()}.
  \item Create three separate loops to iterate through each object in that
    list.
  \item In the first loop, implement logic to determine if a move exists that
    leads to an extra move. Achieve this by comparing the number of seeds that
    would be sown and the distance to the player's store. If equal, that move
    leads to an extra move. Otherwise it does not.
  \item In the second loop, implement logic to determine if a move exists that
    leads to a capture. Achieve this by calculating where the last seed sown
    will land and using this information to determine whether a capture occurs
    at that location.
  \item In the third loop, implement logic to determine the legal move that
    exists with the lowest house number i.e. index.
  \item Find the \texttt{Kalah} class.
  \item Find the \texttt{Play} method.
  \item Within the \texttt{while} loop, modify the assignment to the variable
    \texttt{selection} to read: \texttt{selection = (playerNum == 1)
    ? GameIO.computeMove(board, playerNum, io) : GameIO.getMove(board,
    playerNum, io);}.
\end{enumerate}

\section{Implementing the Change Case}
The total amount of time taken to implement the change case was timed as being 
15 minutes and 58 seconds. During the execution of the aforementioned plan, there were some
\begin{enumerate}
  \item No issues in finding the \texttt{GameConfig} class.
  \item No issues in changing the value of \texttt{VERTICAL\_OUTPUT}.
  \item No issues in finding the \texttt{GameIO} class.
  \item No issues in creating the \texttt{computeMove} method.
  \item No issues in creating a \texttt{List<Pit>} or invoking the required
    methods to obtain a value for it.
  \item No issues in creating the loops.
  \item In the first loop, implement logic to determine if a move exists that
    leads to an extra move. Achieve this by comparing the number of seeds that
    would be sown and the distance to the player's store. If equal, that move
    leads to an extra move. Otherwise it does not.
  \item In the second loop, implement logic to determine if a move exists that
    leads to a capture. Achieve this by calculating where the last seed sown
    will land and using this information to determine whether a capture occurs
    at that location.
  \item In the third loop, implement logic to determine the legal move that
    exists with the lowest house number i.e. index.
  \item No issues in finding the \texttt{Kalah} class.
  \item No issues in finding the \texttt{Play} method.
  \item No issues in modifying the assignment of \texttt{selection}.
\end{enumerate}

\section{Impact of the Change Case}
On a scale from \textit{0.0} to \textit{1.0}, the impact of this change case is
\textit{0.3}. Although the rules of \textit{Kalah} are preserved and the
overall functionality of the game remains largely unchanged, the replacement 
of the second player with a simple AI changes the game from being multi-player 
to being single-player.

\section{Changeability Assessment}
Implementing this change case required a large amount of new code to be
written, and this exposed some gaps in the changeability of the design. This
change case was implemented on top of a previous change case
\cite{a4-change-case}, and in that instance the design showed a good level of 
changeability. However, due to a lack of factorisation of functionality related
to determining different move outcomes (i.e. gaining an extra move or executing
a capture), that aspect of the design did not have good changeability. If
a future developer wishes to change the conditions that lead to an extra move
or a capture they will need to significantly alter the design and doing so will
not be a straightforward process, as I have discovered in implementing this
change case. So, contrary to claims made previously \cite{a3-design}, this
design of Kalah only \textit{partially} lends itself to changeability and could
be improved to be more changeable.

\begin{thebibliography}{00}
  \bibitem{a3-design} M. Eden ``SOFTENG 701 - Assignment 3 - Submission``,
    \textit{Implementing Kalah with a Changeable Object-Oriented Design}
  \bibitem{a4-change-case} M. Eden ``SOFTENG 701 - Assignment 4 - Submission``,
    \textit{Implementing the Change Board Orientation change case}
\end{thebibliography}
\end{document}
